---
title: "Hepatitis A outbreak in Louisville (2017-2019) - Surveillance"
author: "Emmanuelle Dankwa (dankwa@stats.ox.ac.uk)"
date: "08/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Load package
library(dplyr)

# Load functions 
source("surveillance_functions.R")     # All plot functions are written in this file

# Load data 
# Data on general population
dat2 <- readRDS(file = "./surveillance_data/dat2.RDS")                # Full
dat2_yrwk <- readRDS(file = "./surveillance_data/dat2_yrwk.RDS")  #  Summarized (Weekly case counts in the general population)

# Weekly case count among PEH/PWUD (=risk group)
target <- readRDS(file = "./surveillance_data/target.RDS")                 # Full 
target_yrwk <- readRDS(file = "./surveillance_data/target_yrwk.RDS") # Summarized (Vaccination counts by week among PEH/PWUD)

# Vaccination data
vacc_counts_target <-  readRDS(file = "./surveillance_data/vacc_counts_target.RDS") # Summarized (Vaccination counts by week among PEH/PWUD)


```


# Summaries (Table 2)

<!-- The code below reproduces Table 2.  -->


```{r table-2}
## Define PEH/PWUD and recategorize housing status variable

dat2 <- dat2 %>%
mutate(PEHPWUD =  Homelessness == 'Homeless - Shelter/Streets' | Homelessness== 'Unstable Housing'| IVDU=='Yes' | Non.IV.DU=='Yes')%>%
mutate(Homelessness  =  if_else(Homelessness == "Homeless - Shelter/Streets" | Homelessness== 'Unstable Housing', 'Homeless',
          if_else(Homelessness == 'Unknown' | Homelessness == '', 'Unknown', 'Not homeless'))) %>% 
  mutate(Sex = factor(Sex, levels = c('Male', 'Female'))) %>% 
  mutate(Non.IV.DU = factor(Non.IV.DU, levels = c('Yes', 'No', 'Unknown'))) %>% 
  mutate(IVDU = factor(IVDU, levels = c('Yes', 'No', 'Unknown'))) %>% 
  mutate(PEHPWUD = factor(PEHPWUD, levels = c('TRUE', 'FALSE')))



# Generate summaries 
s <- apply(dat2, 2 , generate.summaries)

```



**Table 2:** Summary of detected cases by year, sex, age group, housing status and illicit drug use.


| Variable                          | Number of cases (\%)                                                 | Number of hospitalizations (\%)                                     | Number of deaths (\%)                                               |
|-----------------------------------|----------------------------------------------------------------------|--------------------------------------------------------------------|--------------------------------------------------------------------|
| **Year**                              |                                                                      |                                                                    |                                                                    |
| 2017                              | `r s$MMWRYear$cases[[1]]` (`r   s$MMWRYear$cases_perc[[1]]`)         | `r s$MMWRYear$hosp[[1]]` (`r   s$MMWRYear$hosp_perc[[1]]`)         | `r s$MMWRYear$mort[[1]]` (`r   s$MMWRYear$mort_perc[[1]]`)         |
| 2018                              | `r s$MMWRYear$cases[[2]]` (`r   s$MMWRYear$cases_perc[[2]]`)         | `r s$MMWRYear$hosp[[2]]` (`r   s$MMWRYear$hosp_perc[[2]]`)         | `r s$MMWRYear$mort[[2]]` (`r   s$MMWRYear$mort_perc[[2]]`)         |
| 2019                              | `r s$MMWRYear$cases[[3]]` (`r   s$MMWRYear$cases_perc[[3]]`)         | `r s$MMWRYear$hosp[[3]]` (`r   s$MMWRYear$hosp_perc[[3]]`)         | `r s$MMWRYear$mort[[3]]` (`r   s$MMWRYear$mort_perc[[3]]`)         |
| **Sex**                               |                                                                      |                                                                    |                                                                    |
| Male                              | `r s$Sex$cases[[2]]` (`r   s$Sex$cases_perc[[2]]`)                   | `r s$Sex$hosp[[2]]` (`r   s$Sex$hosp_perc[[2]]`)                   | `r s$Sex$mort[[2]]` (`r   s$Sex$mort_perc[[2]]`)                   |
| Female                            | `r s$Sex$cases[[1]]` (`r   s$Sex$cases_perc[[1]]`)                   | `r s$Sex$hosp[[1]]` (`r   s$Sex$hosp_perc[[1]]`)                   | `r s$Sex$mort[[1]]` (`r   s$Sex$mort_perc[[1]]`)                   |
| **Age (in years)**                    |                                                                      |                                                                    |                                                                    |
| 10-19                             | `r s$Age_group$cases[[1]]` (`r   s$Age_group$cases_perc[[1]]`)       | `r s$Age_group$hosp[[1]]` (`r   s$Age_group$hosp_perc[[1]]`)       | `r s$Age_group$mort[[1]]` (`r   s$Age_group$mort_perc[[1]]`)       |
| 20-29                             | `r s$Age_group$cases[[2]]` (`r   s$Age_group$cases_perc[[2]]`)       | `r s$Age_group$hosp[[2]]` (`r   s$Age_group$hosp_perc[[2]]`)       | `r s$Age_group$mort[[2]]` (`r   s$Age_group$mort_perc[[2]]`)       |
| 30-39                             | `r s$Age_group$cases[[3]]` (`r   s$Age_group$cases_perc[[3]]`)       | `r s$Age_group$hosp[[3]]` (`r   s$Age_group$hosp_perc[[3]]`)       | `r s$Age_group$mort[[3]]` (`r   s$Age_group$mort_perc[[3]]`)       |
| 40-49                             | `r s$Age_group$cases[[4]]` (`r   s$Age_group$cases_perc[[4]]`)       | `r s$Age_group$hosp[[4]]` (`r   s$Age_group$hosp_perc[[4]]`)       | `r s$Age_group$mort[[4]]` (`r   s$Age_group$mort_perc[[4]]`)       |
| 50-59                             | `r s$Age_group$cases[[5]]` (`r   s$Age_group$cases_perc[[5]]`)       | `r s$Age_group$hosp[[5]]` (`r   s$Age_group$hosp_perc[[5]]`)       | `r s$Age_group$mort[[5]]` (`r   s$Age_group$mort_perc[[5]]`)       |
| 60-69                             | `r s$Age_group$cases[[6]]` (`r   s$Age_group$cases_perc[[6]]`)       | `r s$Age_group$hosp[[6]]` (`r   s$Age_group$hosp_perc[[6]]`)       | `r s$Age_group$mort[[6]]` (`r   s$Age_group$mort_perc[[6]]`)       |
| 70+                               | `r s$Age_group$cases[[7]]` (`r   s$Age_group$cases_perc[[7]]`)       | `r s$Age_group$hosp[[7]]` (`r   s$Age_group$hosp_perc[[7]]`)       | `r s$Age_group$mort[[7]]` (`r   s$Age_group$mort_perc[[7]]`)       |
| **Housing   status**                  |                                                                      |                                                                    |                                                                    |
| Homeless                          | `r s$Homelessness$cases[[1]]`   (`r s$Homelessness$cases_perc[[1]]`) | `r s$Homelessness$hosp[[1]]` (`r   s$Homelessness$hosp_perc[[1]]`) | `r s$Homelessness$mort[[1]]` (`r   s$Homelessness$mort_perc[[1]]`) |
| Not   homeless                    | `r s$Homelessness$cases[[2]]`   (`r s$Homelessness$cases_perc[[2]]`) | `r s$Homelessness$hosp[[2]]` (`r   s$Homelessness$hosp_perc[[2]]`) | `r s$Homelessness$mort[[2]]` (`r   s$Homelessness$mort_perc[[2]]`) |
| Unknown                           | `r s$Homelessness$cases[[3]]`   (`r s$Homelessness$cases_perc[[3]]`) | `r s$Homelessness$hosp[[3]]` (`r   s$Homelessness$hosp_perc[[3]]`) | `r s$Homelessness$mort[[3]]` (`r   s$Homelessness$mort_perc[[3]]`) |
| **Illicit drug use, intravenous**     |                                                                      |                                                                    |                                                                    |
| Yes                               | `r s$IVDU$cases[[3]]` (`r   s$IVDU$cases_perc[[3]]`)                 | `r s$IVDU$hosp[[3]]` (`r   s$IVDU$hosp_perc[[3]]`)                 | `r s$IVDU$mort[[3]]` (`r   s$IVDU$mort_perc[[3]]`)                 |
| No                                | `r s$IVDU$cases[[1]]` (`r   s$IVDU$cases_perc[[1]]`)                 | `r s$IVDU$hosp[[1]]` (`r   s$IVDU$hosp_perc[[1]]`)                 | `r s$IVDU$mort[[1]]` (`r   s$IVDU$mort_perc[[1]]`)                 |
| Unknown                           | `r s$IVDU$cases[[2]]` (`r   s$IVDU$cases_perc[[2]]`)                 | `r s$IVDU$hosp[[2]]` (`r   s$IVDU$hosp_perc[[2]]`)                 | `r s$IVDU$mort[[2]]` (`r   s$IVDU$mort_perc[[2]]`)                 |
| **Illicit drug use, non-intravenous** |                                                                      |                                                                    |                                                                    |
| Yes                               | `r s$Non.IV.DU$cases[[3]]` (`r   s$Non.IV.DU$cases_perc[[3]]`)       | `r s$Non.IV.DU$hosp[[3]]` (`r   s$Non.IV.DU$hosp_perc[[3]]`)       | `r s$Non.IV.DU$mort[[3]]` (`r   s$Non.IV.DU$mort_perc[[3]]`)       |
| No                                | `r s$Non.IV.DU$cases[[1]]` (`r   s$Non.IV.DU$cases_perc[[1]]`)       | `r s$Non.IV.DU$hosp[[1]]` (`r   s$Non.IV.DU$hosp_perc[[1]]`)       | `r s$Non.IV.DU$mort[[1]]` (`r   s$Non.IV.DU$mort_perc[[1]]`)       |
| Unknown                           | `r s$Non.IV.DU$cases[[2]]` (`r   s$Non.IV.DU$cases_perc[[2]]`)       | `r s$Non.IV.DU$hosp[[2]]` (`r   s$Non.IV.DU$hosp_perc[[2]]`)       | `r s$Non.IV.DU$mort[[2]]` (`r   s$Non.IV.DU$mort_perc[[2]]`)       |
| **PEH/PWUD^***                        |                                                                      |                                                                    |                                                                    |
| Yes                               | `r s$PEHPWUD$cases[[2]]` (`r   s$PEHPWUD$cases_perc[[2]]`)       | `r s$PEHPWUD$hosp[[2]]` (`r   s$PEHPWUD$hosp_perc[[2]]`)       | `r s$PEHPWUD$mort[[2]]` (`r   s$PEHPWUD$mort_perc[[2]]`)       |
| No                                | `r s$PEHPWUD$cases[[1]]` (`r   s$PEHPWUD$cases_perc[[1]]`)       | `r s$PEHPWUD$hosp[[1]]` (`r   s$PEHPWUD$hosp_perc[[1]]`)       | `r s$PEHPWUD$mort[[1]]` (`r   s$PEHPWUD$mort_perc[[1]]`)       |
| **Total**                             | `r nrow(dat2)` (100\%)                                               | `r sum(dat2$Hospitalization==   'Yes')` (100\%)                    | `r sum(dat2$Mortality== 'Yes')`   (100\%)                          |





# Vaccination and case count plots (Figure 2)

The code in this chunk reproduces Figure 2.

```{r fig.align ="center", fig.cap = "Weekly case and vaccination counts (Figure 2)", fig.height= 15}
# Plot (uncomment the commented lines below and modify directory (argument of "file") to save plot locally)
#png(file = "./figs/for_paper/casecounts_vaccinations_final.png", res = 800, width = 10, height = 14, units = "in")
layout(matrix(c(1,2),2, 1, byrow = FALSE))
plot.by.risk(dat2_yrwk)                                   # Detected cases
plot.vaccinations.comparable.axis(vacc_counts_target)     # Vaccinations
# dev.off()
```

